---
title: "hw04_Join_Cheetsheet"
output: github_document
---

```{r}
library(gapminder)
library(tidyverse)
library(reshape2)
```

###Join, merge, look up

####Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins.

Solution:

Created second data frame with Country and its Capital City.
```{r}
country_capitals <- data.frame(country=c("Australia","Austria","Belgium","China", "Italy", "Croatia", "Egypt", "Greece", "India"), capital=c("CANBERRA","VIENNA","BRUSSELS","BEIJING","ROME","ZAGREB","CAIRO","ATHENS","NEW DELHI"))
knitr::kable(country_capitals)
gapminder
```


##### Left Join

left_join() : Retains the column country in gapminder dataset and appends the capital city(New data) as a new column into the gampminder dataset.
left_join returns data from the second data frame aswell.Please note Column 7 in the below table (i.e. capital city)

```{r}

Left_Join <-left_join(gapminder,country_capitals ,by = "country")
Left_Join %>% filter(country %in% c("Australia","Austria","Belgium","China", "Italy", "Croatia", "Egypt", "Greece", "India"))

```

##### Anti Join

anti_join() : Retains the coutries that are present in gapminder dataset but not in the second data set.
And hence in the below example it only displays 1,596 rows(i.e Gapminder (1704 rows) - Second data set(108 rows))
anti_join does not contain return any column of the second data set.

```{r}
Anti_join<-anti_join(gapminder,country_capitals)
Anti_join
```

##### Semi Join

Semi_join() : Retains the countries that are common in both gapminder dataset and in the second data set.However,it does not return the column present in second data frame,hence the below table contains 6 columns only.

```{r}
Semi_join<-semi_join(gapminder,country_capitals)
Semi_join %>% filter(country %in% c("Australia","Austria","Belgium","China", "Italy", "Croatia", "Egypt", "Greece", "India"))
```

##### Full Join

Full_join() :Retains the countries from both gapminder and the second data set and appends additional information against each country.Updates unavailable data as NA in the last column(i.e..,column 7)

```{r}
Full_join <-full_join(gapminder,country_capitals)
Full_join %>% filter(country %in% c("Australia","Austria","Belgium","China", "Italy", "Croatia", "Egypt", "Greece", "India"))
```

##### Inner Join

Inner_join() :Retains the countries that are common to both gapminder dataset and the  second data set and also, appends a new column for new information against each country.
```{r}
Inner_join <-inner_join(gapminder,country_capitals)
Inner_join %>% filter(country %in% c("Australia","Austria","Belgium","China", "Italy", "Croatia", "Egypt", "Greece", "India"))
```




