---
title: "https://github.com/jmurthy12/STAT545-hw02-murthy-janani"
output: github_document
---
Bring rectangular data in


```{r}
library(gapminder)
library(tidyverse)
```

###Smell test the data
Explore the gapminder object:

1)Is it a data.frame, a matrix, a vector, a list?

Answer : Gapminder object is a data frame .

```{r}
gapminder
```
2)What's its class?

Answer  : Class is Tibble.
```{r}
class(gapminder)
```

OR 
An alternative way is to find the number of rows and columns is mentioned below :
```{r}
str(gapminder)
```

3)How many variables/columns?
There are 6 variables/coulmns .
The Variables are : country,continent,year,lifeExp,pop,gdpPercap
```{r}
gapminder
```

4)How many rows/observations?
Answer : There are 1,1704 rows/observartions



5)Can you get these facts about "extent" or "size" in more than one way? Can you imagine different functions being useful in different contexts?
```{r}
dim(gapminder)
nrow(gapminder)
ncol(gapminder)
length(gapminder$country)
```

6)What data type is each variable?
```{r}
class(gapminder$country)
class(gapminder$continent)
class(gapminder$year)
class(gapminder$lifeExp)
class(gapminder$pop)
class(gapminder$gdpPercap)
```


###Explore individual variables : 

1)Pick at least one categorical variable and at least one quantitative variable to explore.

Answer : We extract the continent variable as an example for a categorical variable and the lifeExp as continuous variable
ContinentVar <- gapminder$continent
lifeExpVar <- gapminder$lifeExp
```{r}
ContinentVar <- gapminder$continent
lifeExpVar <- gapminder$lifeExp
```



2)What are possible values (or range, whichever is appropriate) of each variable?
continentTable <- table(ContinentVar)
    What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
    Feel free to use summary stats, tables, figures. We’re NOT expecting high production value (yet).
    
    
```{r}
continentTable <- table(ContinentVar)
continentTable
hist(continentTable)
range(lifeExpVar)
mean(lifeExpVar)
sd(lifeExpVar)

```
```{r}
gap_with_colors <-
data.frame(gapminder,
cc = I(country_colors[match(gapminder$country,
names(country_colors))]))
```


##Explore various plot types

See the ggplot2 tutorial, which also uses the gapminder data, for ideas.

Make a few plots, probably of the same variable you chose to characterize numerically. Try to explore more than one plot type. Just as an example of what I mean:

    A scatterplot of two quantitative variables.
    A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.
    A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries.


```{r}
library(gapminder)
library("ggplot2")
ggplot(gapminder, aes(x = lifeExp, y = gdpPercap)) +
  geom_point()
ggplot(gapminder, aes(x = ))

```






