---
title: "https://github.com/jmurthy12/STAT545-hw02-murthy-janani"
output: github_document
---
# Exploring Gapminder!
### Bring rectangular data in

```{r}
library(gapminder)
library(tidyverse)
```


### Smell test the data
Explore the gapminder object:

1)Is it a data.frame, a matrix, a vector, a list?

Answer : List.Gapminder is a dataframe and a List.

```{r}
is.data.frame(gapminder)
is.list(gapminder)
typeof(gapminder)
```

2)What's its class?

Answer  : Class is dataframe and also a tibble.
```{r}
class(gapminder)
gapminder
```

OR 

An alternative way is to find the number of rows and columns as mentioned below :
```{r}
str(gapminder)
```

3)How many variables/columns?

Answer :There are 6 variables/coulmns .
The Variables are : country,continent,year,lifeExp,pop and gdpPercap
```{r}
gapminder
```

4)How many rows/observations?

Answer : There are 1,1704 rows/observartions



5)Can you get these facts about "extent" or "size" in more than one way? Can you imagine different functions being useful in different contexts?

Answer :  Various methods/functions to find the "extent" or "size" are described below :

```{r}
dim(gapminder)
nrow(gapminder)
ncol(gapminder)
head(gapminder) #gives the number of columns and data types for each column
length(gapminder) # gives the number of columns
```

6)What data type is each variable?
```{r}
class(gapminder$country)
class(gapminder$continent)
class(gapminder$year)
class(gapminder$lifeExp)
class(gapminder$pop)
class(gapminder$gdpPercap)
```

OR, 

alternatively we could use the str() command

```{r}
str(gapminder) # gives the consolidated data type for each variable/column.
```

```{r}
summary(gapminder)
```


### Explore individual variables : 

1)Pick at least one categorical variable and at least one quantitative variable to explore.

Answer :
Categorical variable : continent  
Quantitative variable : life expectancy(lifeExp)

```{r}
ContinentVariable <- gapminder$continent
lifeExpVariable <- gapminder$lifeExp
```



2)What are possible values (or range, whichever is appropriate) of each variable?
    What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
    Feel free to use summary stats, tables, figures. We’re NOT expecting high production value (yet).
    
    
```{r}
continentTable <- table(ContinentVariable)
continentTable
hist(continentTable) #Distribution of Continent (Similar to frequency distribution)
range(lifeExpVariable) # Range of Life Expectancy
hist(lifeExpVariable) #Distribution of ife Expectency
mean(lifeExpVariable) # Mean/Average of ife Expectancy
sd(lifeExpVariable) # Standard Deviation of Life Expectancy

```

Exploring continents and Life expectancy

```{r}
	gapminder %>%
	  filter(continent== "Asia") %>%
	  group_by(year) %>%
	  summarise (minexp = min(lifeExp),
             maxexp = max(lifeExp)) # displays the minimum and maximum Life expectancy
```

```{r}
unique(gapminder$continent) #gives the distinct countries within continent Variable set
```



## Explore various plot types

See the ggplot2 tutorial, which also uses the gapminder data, for ideas.

##### Make a few plots, probably of the same variable you chose to characterize numerically. Try to explore more than one plot type. Just as an example of what I mean:

   A scatterplot of two quantitative variables.
   
   
   Quantitative variable 1 : Life Expectancy 
   
   Quantitative variable 2 : Population
    
```{r}
ggplot(gapminder, 
      aes(x=lifeExp, y=pop)) +
      geom_point(size = 2, color = "indianred4")
```
    
    A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.
    
    Densityplot of  quantitative variable: LifeExpectancy
    
```{r}
ggplot(gapminder, aes(x=lifeExp)) + geom_density()
```

A plot of one quantitative variable and one categorical


Answer :  Box plot of quantitative variable : LifeExpectancy vs categorical variable  : Continent

```{r}
ggplot(gapminder,
      aes(x = continent, y = lifeExp)) +
      geom_boxplot(colour = "azure4", fill = "cyan", alpha = 0.5)
```

     Maybe boxplots for several continents or countries.


```{r}

ggplot(gapminder,
      aes(x=continent, y=lifeExp)) +
      geom_boxplot(colour = "blue2", fill = "bisque2", alpha = 0.5) +
      ggtitle("Continents") +
      labs(x = "Country", y = "Lifeexpectancy")
```

```{r}

ggplot(gapminder, aes(x = lifeExp, y = gdpPercap)) +
  geom_point()
ggplot(gapminder, aes(x =lifeExp, y = year))+
  geom_point()

```


Use filter(), select() and %>%
Use filter() to create data subsets that you want to plot.

Practice piping together filter() and select(). Possibly even piping into ggplot().


Answer : 
All the above functionaities mentioned above are excuted in the chunk below.


Scatter Plot describing the LifeExpectency by year for each country within the Asian Continent.

```{r}
gapminder %>%
  filter(continent=="Asia") %>% 
  select(country, year, lifeExp) %>% 
  ggplot(aes(x=year, y=lifeExp)) %>% 
  + geom_point(aes(color=country))
```


### But I want to do more! 
For people who want to take things further.

Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?

 
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

##### Similar excercise was done in class cm006 and the explaination was provided in the same class.


Explaination :The above chunk only displays countries(afghanistan and Rwanda)/rows for which the year value/columns are unique.
Repetitions are omitted.Hence it displays 12 rows only instead of 24 rows.

Solution : To display all the data for country Rwanda nd Afghanistan, use the below :

```{r}
filter(gapminder, country == "Rwanda" | country == "Afghanistan") # As discussed in class cm006
```

OR

```{r}

gapminder %>%
  filter(country %in% c("Rwanda", "Afghanistan"))  # As discussed in class cm006
```



Present numerical tables in a more attractive form, such as using knitr::kable().

Presenting the above example using knitr::kable() 

```{r}
knitr::kable(filter(gapminder, country == "Rwanda" | country == "Afghanistan"))

```




