---
title: "https://github.com/jmurthy12/STAT545-hw02-murthy-janani"
output: github_document
---
Bring rectangular data in


```{r}
library(gapminder)
library(tidyverse)
```

Smell test the data
Explore the gapminder object:

1)Is it a data.frame, a matrix, a vector, a list?
Answer : Gapminder object is a data frame .

```{r}
gapminder
```
2)What’s its class?
Answer  : Class is Tibble.
```{r}
class(gapminder)
```


3)How many variables/columns?
There are 6 variables/coulmns .
The Variables are : country,continent,year,lifeExp,pop,gdpPercap
```{r}
gapminder
```

4)How many rows/observations?
Answer : There are 1,1704 rows/observartions

5)Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
```{r}
dim(gapminder)
nrow(gapminder)
ncol(gapminder)
length(gapminder$country)
```

6)What data type is each variable?
```{r}
class(gapminder$country)
class(gapminder$continent)
class(gapminder$year)
class(gapminder$lifeExp)
class(gapminder$pop)
class(gapminder$gdpPercap)
```





