---
title: "hw03_gapminder.Rmd"
output: github_document
---

Loading the libraries
```{r}

library(gapminder)
library(tidyverse)
library(knitr)
```



#### Query 1 : Get the maximum and minimum of GDP per capita for all continents


Solution : 
```{r}
minmaxGDP<- gapminder %>%
  group_by(continent) %>%
  summarise(max_gdp = max(gdpPercap), min_gdp = min(gdpPercap))
kable(minmaxGDP)

```
```{r}
ggplot(minmaxGDP,aes(x=continent,y=max_gdp)) + geom_bar(stat="identity")
ggplot(minmaxGDP,aes(x=continent,y=min_gdp)) + geom_bar(stat="identity")

```

#### Query 2 : Look at the spread of GDP per capita within the continents.


Solution :


```{r}
spread <- gapminder %>%
  group_by(continent) %>% 
  summarize(mean_gdp = mean(gdpPercap), 
            sd_gdp = sd(gdpPercap), 
            max_gdp = max(gdpPercap),
            min_gdp = min(gdpPercap))
kable(spread) 
```
A box plot showing the spread of GDP per capita within the continents

```{r}
ggplot(gapminder, aes(continent, gdpPercap)) +
  geom_boxplot(aes(color=continent))
```

Mean value for gdpPerCap for each country  over all years is calculated and a histogram is plotted to indicate the gdpPercap ,continentwise.

```{r}
spread <- gapminder %>%
  group_by(country,continent) %>%
  summarize(gdpAverage=mean(gdpPercap))
spread <- arrange(spread,continent)

knitr::kable((spread)) 
ggplot(spread,aes(x=gdpAverage,fill ="purple"))  + facet_wrap(~ continent) + geom_histogram(colour="cyan")
```
```{r}
gapminder %>%
  group_by(continent) %>%
  ggplot(aes(year,gdpPercap,alpha=.2)) + 
  geom_point(size = 1.5, color = "indianred4") +
  facet_wrap(~continent)
```


#### Query 3 : Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.


Solution : 
Weighing Life Expectancy by population is tried below :

```{r}
WeightedLifeExp <- gapminder %>% 
  group_by(year) %>% 
  summarize(WeightedLifeExpByPop = weighted.mean(lifeExp,pop))
knitr::kable(WeightedLifeExp)
```


#### Query 4 How is life expectancy changing over time on different continents?


Solution : 
```{r}
ggplot(gapminder, aes(year, lifeExp)) +
    facet_wrap(~ continent) +
    geom_point()+ geom_smooth(se=F)
```

Over the years,in general, life expectency has increased for all continents.Life expectency for the continents :Asia and Americas , shows an upward trend compared to those of Europe and Oceania show a gradual increase in Life Expectancy.
While in Africa, there is gradual increase in LifeExpectancy from 1950 - 1985,but the same almost stabilises between the years post 1985.


#### Query 5 : Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.


Solution : 
```{r}
Benchmark <- mean(gapminder$lifeExp)
life_threshold <-gapminder %>% 
  group_by(year,continent) %>% 
  mutate(size=c(1,0)[(lifeExp<Benchmark) + 1]) %>%
  mutate(size_con=c(1,0)[(lifeExp<0) + 1]) %>% 
  summarise(percentage = sum(size)/sum(size_con)*100) %>% 
  arrange(continent)
summary(life_threshold$percentage)
knitr::kable((life_threshold))
ggplot(life_threshold,aes(year,percentage,color=continent)) + geom_line()
```

#### Query 6: Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.


Solution : 
Countries Selected : India and China

```{r}

India_China <- gapminder %>%
  filter(country %in% c("India", "China")) # filtering for only India and China


India_China %>%
  ggplot(aes(x = year))+
  geom_line(aes(y = pop), colour = "orange") +
  geom_line(aes(y = lifeExp), colour = "indianred4") +
  scale_y_continuous(sec.axis = sec_axis(~.*1, name = "LifeExpectancy")) +
  facet_wrap(~country)# plotting graphs to compare pop and GDP
```

India and China show an upward trend for Population growth,while the LifeExpectancy stagnates.
```{r}
# filtering for only India and China
India_China <- gapminder %>%
  filter(country %in% c("India", "China")) 

# plotting using 2 different y axis in order to compare pop and GDP
India_China %>%
  ggplot(aes(x = year))+
  geom_line(aes(y = gdpPercap), colour = "orange") +
  geom_line(aes(y = pop), colour = "indianred4") +
  scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Population")) +
  facet_wrap(~country)
```

In the above graph aswell,India and China show an upward trend for Population growth,while the GdpPer captica stagnates.
